The inefficacy of superficial similarities for improving instructor-student relationships

### 2021-04-26 outputs/paper/paper.Rmd
shiny complete, train/test complete, first round of content edits done.

---
title: "The inefficacy of superficial similarities for improving instructor-student relationships"
author: "Amy Farrow"
date: "April 19th, 2021"
header-includes:
  \usepackage{dcolumn}
  \usepackage{rotating}
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "Replicating ‘Taking It to the Next Level’"
abstract: "This paper replicates the 2019 article 'Taking It to the Next Level' [@citeoriginal], which evaluates an intervention to improve college instructor-student relationship. The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity or instructor-student relationship, grades, or re-enrollment. The goal of this intervention was to improve student performance and persistence, and while the scalability and affordability are desirable, there are no results in any of the targeted measures. These results are consistent with the original paper.\\par \\textbf{Keywords:} instructor-student relationship, college, replication study"
thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
bibliography: references.bib
---

@@

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in an RProject
library(Hmisc) # for correlation matrix
library(readstata13) # for stata files
library(finalfit) # for summary table
library(stargazer) # for model table
library(MASS) # for ordinal logistic model
library(cowplot) # for putting plots side by side
library(ltm) # for checking scale validity
library(DiagrammeR) # for causal diagram

@@

TODO: Shiny. Recreate Figure \@ref(fig:continuousoutcomes)(add graphs for standardized grade and persistence). Make an interactive element where you can filter the graph for a subset of the data -- specific racial s_race, firstgen/not, specific year of study.

@@

Despite the growing importance of college degrees for career success, college completion rates are a continuing problem. The causes are complex: factors that affect student persistence include student characteristics, organizational factors, peer environment, and student experiences. Interventions to improve student retention can target any of these aspects.

@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college retention and performance. In this field experiment, they evaluated the relationships between perceived similarity, instructor-student relationships (ISRs), and measures of student success. Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result. The experiment consisted of a randomized controlled trial where some undergraduate students were informed of similarities they shared with their instructor, while others were not. Student and instructor perceptions of similarity and the instructor-student relationship were measured through surveys, and student performance measures were collected from school records.

This paper replicates @citeoriginal's original analysis, using anonymized data provided by the authors. First, @citeoriginal's methods are discussed, and the measures and scales are evaluated. Second, demographics of instructor and student populations are explored, similarity of treatment and control groups is established, outcomes of interest are compared for treatment and control groups, and missing data is evaluated. Third, @citeoriginal's models are replicated, and additional models are explored. Linear, logistic, and ordinal logistic models are replicated to predict outcomes including ISR perception, grade, and persistence.  Further replicated exploratory models consider the relationship between similarity and instructor-student relationship, as well as between instructor-student relationship and student outcomes. Additional models use different controls--specifically, variables indicating shared racial and gender self-identification. Finally,

These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grades, or persistence. These conclusions stand when matching instructor-student traits are considered and when vulnerable student groups are evaluated separately from the wider student population. The lack of significant effects reflects the complicated task of increasing perceived similarity, hopefully to improve instructor relationships, and to then improve student outcomes, especially in a college context that may make instructor-student similarity and connection challenging.

@@

Beyond the wide range of research that addresses similarity, ISRs, and student outcomes, @citeoriginal was directly inspired by two previous works in middle and high schools. @Gehlbach_Brinkworth_Harris_2012's study of middle-school students found that students' perceived similarity to their teachers positively correlated with improvements in teacher-student relationship, and that changes in teacher-student relationships are associated with changes in student perceived self-efficacy, but not scholastic performance. Using a randomized controlled trial and surveys, @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015 found that when high school students were presented with similarities they shared with their teachers, they subsequently perceived greater similarity with their teachers. When teachers were presented with similarities, they perceived better relationships with those students, and those students subsequently performed better academically. For students, however, awareness of similarities did not effectively improve their perception of the student-teacher relationship.

@@

@citeoriginal conducted a randomized controlled trial to assess the impact of awareness of instructor-student similarities on similarity perception, instructor-student relationship, course grade, and re-enrollment.

@@

The study took place in the 2017 spring term at a large Californian University. The study included 120 instructors and their 2,749 students. The instructors participated in the study based on interest and a gift-card incentive, and their students were invited to participate unincentivized. Students were only enrolled in the study for one class, in the event that they were taking classes with multiple participating instructors. The initial sample of 147 instructors was a convenience sample: the study was advertised, interested instructors signed up and were enrolled if their course met the study requirements. 145 instructors consented and took the initial survey, resulting in an corresponding sample of 3,352 students. Once errors of administration, missing responses, and inadequate time spent on surveys were considered, the sample consisted of only 119 instructors and 2,273 students.

@@

All students participated in a survey immediately following the treatment or the placebo. They were surveyed again at the end of the course. Instructors were surveyed only at the end of the course.

@@

Some are extracted from the student survey at the beginning of the term:

  1. Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5. These responses were averaged to create a student perceived similarity scale.
  2. Immediately after the treatment or placebo, students answered seven questions about their anticipated instructor-student relationship, on scales of 1 to 5. These responses were averaged to create a student anticipated ISR scale.

Others are extracted from the student survey at the end of the term:

  3. At the end of the term, students answered the student perceived similarity scale questions again.
  4. At the end of the term, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5. These responses were averaged to create a student perceived ISR scale.

Others come from the instructor survey at the end of the term:

  5. Instructor similarity perception: At the end of the term, instructors answered only one question about similarity with the student, on a scale of 1 to 5.
  6. Instructor perceived ISR scale: At the end of the term, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5. These responses were averaged to create an ISR scale.
  7. Final grade: Instructors were asked to report the student's grade on their final exam, paper, or project.

Finally, some are extracted from the university's internal records:

  8. The grade that the student received in the course.
  9. The student's course grade, standardized against other grades in the course.
  10. Persistence: The student's status as of Fall term 2017: not enrolled or enrolled.

@@

A scale is externally reliable if it gives the same results across different testings [@Coolican_2014]. Test-retest reliability, where a scale is given to a group at different times, is a common way to assess this [@Coolican_2014]. In our data, there is a retest, but interactions between instructors and students may reasonably be expected to affect the second testing, rendering the data unuseful for assessing test-retest reliability.

@@

# Put it all in a nice table.
demotab %>%
  knitr::kable(caption = "Student covariates for treatment and control groups",
               booktabs = TRUE, linesep = "",
               ) %>%
  column_spec(5, width = "5em") %>%
  column_spec(6, width = "5em") %>%
  column_spec(7, width = "5em") %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
                            stripe_index =c(2:3, 13:14, 20),
                            stripe_color = "#f6f0fb")

@@

There are differences between the students and instructors. Notably, the student sample is 62.7% female, while the instructor sample is 78.2% female. The student sample is 19.8% White, while the instructor sample is 38.7% White. The student sample is 47% Hispanic, while the instructor sample is only 3.4%. Finally, the student sample is 9.8% multiracial, but the instructor sample is 52.1% multiracial. Some of these differences may be due to self-identification: while one person who is Hispanic may select only Hispanic, another may select, for example, Hispanic as well as White, and be classed differently as a result. For this reason, it is very difficult to draw conclusions. Additionally, students are 43.5% first-generation college students, while only 25.2% of instructors are.

In light of the different percentages in student and instructor racial self-identification, Figure \@ref(fig:racegraph) shows the percentages of each group that selected a given racial identity, as well as the percentage that selected more than one of these categories. More students identified as Hispanic, while more instructors identified as White and multi-racial.

@@

```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(5,3), fig.cap="Histogram illustrating class sizes", out.width="60%", fig.align="center"}
expdata %>%
  dplyr::select(teacherid, n_course) %>%
  unique() %>%
  ggplot(aes(x = n_course)) +
  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
  labs(x = "Class size", y = "Frequency", title = "Some class sizes are very large") +
  theme_minimal()
```

## Outcomes of interest

@@

129 eligible instructors completed the first survey, but 6 of those instructors did not complete the second survey. Corresponding to the remaining 123 instructors, 2,801 students completed the first survey (1,392 in the treatment group, 1,409 in the control group), but 682 of those students did not complete the second survey (328 in the treatment group, 354 in the control group). The attrition in the treatment and control groups was comparable (24% in the treatment group, 25% in the control group).

@@

From the 36,838 observations, or potential units of study (corresponding to all undergraduate student records of the university), only 2,273 were used in analysis. Units of study were excluded because the instructor did not participate in the study (33,486); the student did not consent (123); the student did not spend more than a second reading the consent page (1); the initial student survey, initial instructor survey, or end-of-term instructor survey was not complete (428); the instructor mistakenly administered the wrong survey (30); the course was online (20); the course was for graduate students (50); or a participant did not spend more than ten seconds reading and answering a page with five or more questions (427). The ten-second time limit is somewhat arbitrary, and a different choice could lead to different study results.

>>>>

---
title: "The inefficacy of superficial similarities for improving instructor-student relationships"
author: "Amy Farrow"
date: "April 19th, 2021"
header-includes:
  \usepackage{dcolumn}
  \usepackage{float}
  \usepackage{rotating}
  \usepackage{pdflscape}
  \usepackage{booktabs}
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "Replicating ‘Taking It to the Next Level’"
abstract: "This paper replicates the 2019 article 'Taking It to the Next Level' [@citeoriginal], which evaluates an intervention to improve college instructor-student relationships. The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity, or instructor-student relationship. The goal of this intervention was to improve student performance and persistence, and while the scalability and affordability are desirable, there are no measurable effects on any of the student outcome measures, including grades and reenrollment. These results are consistent with the original paper.\\par \\textbf{Keywords:} replication study, instructor-student relationships, college, similarity, college persistence"
thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
bibliography: references.bib
---

@@

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in an RProject
library(Hmisc) # for correlation matrix
library(readstata13) # for stata files
library(finalfit) # for summary table
library(stargazer) # for model table
library(MASS) # for ordinal logistic model
library(cowplot) # for putting plots side by side
library(ltm) # for checking scale validity
library(DiagrammeR) # for causal diagram
library(olsrr)
library(tidymodels) # for train test

@@

@@

Despite the growing importance of college degrees for career success, college completion rates are a continuing problem. The causes are complex: factors that affect student persistence include student characteristics, organizational factors, peer environment, and student experiences. Interventions to improve student persistence can target any of these aspects.

@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college student persistence and performance. In this field experiment, they evaluated the links between perceived similarity, instructor-student relationships (ISRs), and student success. Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive impact on measurable student outcomes. The experiment consisted of a randomized controlled trial where some undergraduate students were informed of similarities they shared with their instructor, while others were not. Student and instructor perceptions of similarity and the instructor-student relationship were measured through surveys, and student performance measures were collected from school records.

This paper replicates @citeoriginal's original analysis, using anonymized data provided by the authors. First, @citeoriginal's methods are discussed, and the measures and scales are evaluated. Second, demographics of instructor and student populations are explored, similarity of treatment and control groups is established, outcomes of interest are compared for treatment and control groups, and missing data is evaluated. Third, @citeoriginal's models are replicated, and additional models are explored. Linear, logistic, and ordinal logistic models are replicated to predict outcomes including ISR perception, grade, and persistence.  Further replicated exploratory models consider the relationship between similarity and instructor-student relationship, as well as between instructor-student relationship and student outcomes. Additional models use different controls--specifically, variables indicating shared racial and gender self-identification.

These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grades, or persistence. These conclusions stand when matching instructor-student traits are considered and when vulnerable student groups are evaluated separately from the wider student population. The lack of significant effects reflects the complicated multi-stage task of increasing perceived similarity, to improve instructor relationships, and to then improve student outcomes, especially in a college context that may make instructor-student similarity and connection challenging.

@@

Beyond the wide range of research that addresses similarity, ISRs, and student outcomes, 'Taking It to the Next Level' was directly inspired by two previous works in middle and high schools. @Gehlbach_Brinkworth_Harris_2012's study of middle-school students found that students' perceived similarity to their teachers positively correlated with improvements in teacher-student relationship, and that changes in teacher-student relationships were associated with changes in student perceived self-efficacy, but not scholastic performance. Using a randomized controlled trial and surveys, @Gehlbach_Brinkworth_King_Hsu_McIntyre_Rogers_2015 found that when high school students were presented with similarities they shared with their teachers, they subsequently perceived greater similarity with their teachers. When teachers were presented with similarities, they perceived better relationships with those students, and those students subsequently performed better academically. For students, however, awareness of similarities did not effectively improve their perception of the student-teacher relationship.

@@

@citeoriginal conducted a randomized controlled trial, with surveys, to assess the impact of awareness of instructor-student similarities on similarity perception, instructor-student relationship perception, grades, and reenrollment.

@@

The study took place in the 2017 spring term at a large Californian University. The study included 148 instructors and their 3,355 students. The instructors participated in the study based on interest and a gift-card incentive, and their students were invited to participate unincentivized. Students were only enrolled in the study for one class, in the event that they were taking classes with multiple participating instructors. The initial sample of 148 instructors was a convenience sample: the study was advertised, interested instructors signed up and were enrolled if their course met the study requirements. 145 instructors consented and took the initial survey, resulting in an corresponding sample of 3,352 students. Once errors of administration, missing responses, and inadequate time spent on surveys were considered, the sample consisted of only 119 instructors and 2,273 students.

@@

At the beginning of the term, the instructors were surveyed, to collect potential similarities. Then, students were surveyed, to collect potential similarities. Immediately, students were either given the treatment or the placebo, and then given further survey questions that asked about similarity and ISR. Students were surveyed again at the end of the term, regarding similarity and ISR. Instructors were surveyed about similarity and ISR only at the end of the term.

@@

  1. Immediately after the treatment or placebo, students answered six questions about their perceived similarity to the instructor, on scales of 1 to 5. These responses were averaged to create a *student perceived similarity scale*.
  2. Immediately after the treatment or placebo, students answered seven questions about their anticipated instructor-student relationship, on scales of 1 to 5. These responses were averaged to create a *student anticipated ISR scale*.
  3. At the end of the term, students answered the *student perceived similarity scale* questions again.
  4. At the end of the term, students answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5. These responses were averaged to create a *student perceived ISR scale*.
  5. At the end of the term, instructors answered only one question about *instructor perceived similarity* with the student, on a scale of 1 to 5.
  6. At the end of the term, instructors answered seven questions about their perception of the instructor-student relationship, on scales of 1 to 5. These responses were averaged to create an *instructor ISR perception* scale.
  7. Instructors were asked to report the student's *grade on their final exam*, paper, or project.
  8. The *grade that the student received in the course* was pulled from internal records.
  9. The student's course grade was *standardized* against other grades in the course.
  10. The student's *status as of Fall term 2017*, either enrolled or not, was pulled from internal records.

@@

A scale is externally reliable if it gives the same results across different testings [@Coolican_2014]. Test-retest reliability, where a scale is given to a group at different times, is a common way to assess this [@Coolican_2014]. In this data, there is a retest, but interactions between instructors and students may reasonably be expected to affect the second testing, rendering the data unuseful for assessing test-retest reliability.

@@

# Put it all in a nice table.
demotab %>%
  knitr::kable(caption = "Student covariates for treatment and control groups",
               booktabs = TRUE, linesep = "",
               ) %>%
  column_spec(5, width = "5em") %>%
  column_spec(6, width = "5em") %>%
  column_spec(7, width = "5em") %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "striped"),
                            stripe_index =c(2:3, 13:14, 20),
                            stripe_color = "#f6f0fb")
```

@@

There are differences between the students and instructors. Notably, the student sample is 62.7% female, while the instructor sample is 78.2% female. The student sample is 19.8% White, while the instructor sample is 38.7% White. The student sample is 47% Hispanic, while the instructor sample is only 3.4%. Finally, the student sample is 9.8% multiracial, but the instructor sample is 52.1% multiracial. Some of these differences may be due to self-identification: while one person who is Hispanic may select only Hispanic, another may select, for example, Hispanic as well as White, and be classified differently as a result. For this reason, it is very difficult to draw conclusions. Additionally, students are 43.5% first-generation college students, while only 25.2% of instructors are.

In light of the different percentages in student and instructor racial self-identification, Figure \@ref(fig:racegraph) shows the percentages of each group that selected a given racial identity, as well as the percentage that selected more than one of these categories. More students identified as Hispanic, while more instructors identified as White and selected multiple options.

@@

```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(9,3), fig.cap="Histogram illustrating class sizes", out.width="100%", fig.align="center"}
n_instructor <- expdata %>%
  dplyr::select(teacherid, n_course) %>%
  unique()

n_by_instructor <- n_instructor %>%
  ggplot(aes(x = n_course)) +
  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
  geom_vline(xintercept = mean(n_instructor %>% pull(n_course), na.rm = TRUE),
             col = "#868cf7",
             lwd = .5) +
  labs(x = "Class size", y = "Number of classes") +
  theme_minimal()

n_student <- expdata %>%
  dplyr::select(id, n_course) %>%
  unique()

n_by_student <- n_student %>%
  ggplot(aes(x = n_course)) +
  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
  geom_vline(xintercept = mean(n_student %>% pull(n_course), na.rm = TRUE),
             col = "#868cf7",
             lwd = .5) +
  labs(x = "Class size", y = "Number of students in study") +
  theme_minimal()

combined <- plot_grid(n_by_instructor, n_by_student,
          ncol = 2)

title <- ggdraw() + 
  draw_label(
    "Some classes are very large (means marked)",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )

plot_grid(title,
          combined,
          legend,
          ncol = 1,
          rel_heights = c(.1, 2))
```

## Outcomes of interest [^1]

[^1]: Interactive graphs and tables showing outcomes of interest are available [here](https://amycfarrow.shinyapps.io/takingittothenextlevel/).

@@

140 eligible instructors completed the first survey, and 2,808 of their students completed the first survey. 21 of those instructors did not complete the second survey. 2,808 students completed the first survey (1,396 in the treatment group, 1,412 in the control group), but 755 of those students did not complete the second survey (364 in the treatment group, 391 in the control group). The attrition in the treatment and control groups was comparable (26.1% in the treatment group, 27.7% in the control group).

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many instructors participated? 148
isrdata %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# How many students could theoretically participate, based on their instructor participating? 3355
isrdata %>%
  filter(!is.na(teacherid)) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many instructors consented to the first survey? 145
isrdata %>%
  filter(t1_consent == 1) %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# How many students does that correspond to? 3352
isrdata %>%
  filter(t1_consent == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# How many of those students consented to survey 1? 3229
isrdata %>%
  filter(t1_consent == 1, s1_consent == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# So total participants in the study: 145 instructors and 3229 students.
# Now reasons for removal from analysis.
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# If we consider non-eligibility, how many instructors are left? 140
isrdata %>%
  filter(t1_consent == 1 & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students") %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# And students? 3112
isrdata %>%
  filter(t1_consent == 1 & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# Check for instructors completing the first survey. 140
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t1_complete == 1) %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# The number of students that corresponds to. 3112
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# Then, how many students completed the first survey? 2808
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

## Treatment: 1396
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & treatment == "Treatment") %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

## Control 1412
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1 
         & treatment == "Control") %>%
  dplyr::select(id) %>%
  unique() %>%
  count()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many instructors completed the second survey? 119
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t1_complete == 1
         & t2_complete == 1) %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# How many students does that correspond to? 2701
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# Students that completed the second survey: 2053
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & s2_complete) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# Treatment 1032
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & s2_complete
         & treatment == "Treatment") %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# Control 1021
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & s2_complete 
         & treatment == "Control") %>%
  dplyr::select(id) %>%
  unique() %>%
  count()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many teachers spent enough time on the survey? 119.
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t1_complete == 1
         & t2_complete == 1
         & t1_timer_consent_3 > 1
         & s1_timer_consent_3 > 1
         & t1_timer_gtky1_3 > 10
         & t1_timer_gtky2_3 > 10
         & t1_timer_gtky3_3 > 10
         & t1_timer_demos1_3 > 10
         & t1_timer_demos2_3 > 10
         & t2_timer > 10) %>%
  dplyr::select(teacherid) %>%
  unique() %>%
  count()

# Students that corresponds to: 2560
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & t1_timer_consent_3 > 1
         & s1_timer_consent_3 > 1
         & t1_timer_gtky1_3 > 10
         & t1_timer_gtky2_3 > 10
         & t1_timer_gtky3_3 > 10
         & t1_timer_demos1_3 > 10
         & t1_timer_demos2_3 > 10
         & t2_timer > 10) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()

# Students that spent enough time on the surveys: 2273
isrdata %>%
  filter(t1_consent == 1 
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & s1_consent == 1
         & t1_complete == 1
         & s1_fullsurvey == 1
         & t2_complete == 1
         & t1_timer_consent_3 > 1
         & s1_timer_consent_3 > 1
         & t1_timer_gtky1_3 > 10
         & t1_timer_gtky2_3 > 10
         & t1_timer_gtky3_3 > 10
         & t1_timer_demos1_3 > 10
         & t1_timer_demos2_3 > 10
         & t2_timer > 10
         & s1_timer_gtky1_3 > 10
         & s1_timer_gtky2_3 > 10
         & s1_timer_gtky3_3 > 10
         & s1_timerpre1_3 > 10
         & s1_timerpre2_3 > 10
         & (s1_timercontrolfeedback_3 > 10 | is.na(s1_timercontrolfeedback_3))
         & (s1_timercontrolresponse_3 > 10 | is.na(s1_timercontrolresponse_3))
         & (s1_timertreatmentfeedba_3 > 10 | is.na(s1_timertreatmentfeedba_3))
         & (s1_timertreatmentrespon_3 > 10 | is.na(s1_timertreatmentrespon_3))
         & s1_timerpost1_3 > 10
         & s1_timerpost2_3 > 10
         & s1_timerdemo1_3 > 10
         & (s2_timer1a_3 > 10 | is.na(s2_timer1a_3))
         & (s2_timer1b_3 > 10 | is.na(s2_timer1b_3))
         & (s2_timer2a_3 > 10 | is.na(s2_timer2a_3))
         & (s2_timer3a_3 > 10 | is.na(s2_timer3a_3))
         & (s2_timer3b_3 > 10 | is.na(s2_timer3b_3))
         & (s2_timer3c_3 > 10 | is.na(s2_timer3c_3))
         & (s2_timer4_3 > 10 | is.na(s2_timer4_3))) %>%
  dplyr::select(id) %>%
  unique() %>%
  count()
```

@@

From the 36,838 observations, or potential units of study (corresponding to all undergraduate student records of the university), only 2,273 were used in analysis. Units of study were excluded because the instructor did not participate in the study (33,483); the instructor or course wasn't actually eligible (96); the instructor administered the wrong survey (31); the student did not consent  (116); the student did not complete the first survey (304); the instructor did not complete the second survey (107); or one participant did not spend at more than a second reading the consent page and more than ten seconds reading each page with five questions or more (428). The time limits are somewhat arbitrary, and different choices could have different analysis results.

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many students were discarded because the instructor did not participate? 33175
all_students <- isrdata %>%
  unique() %>%
  count() %>%
  pull()

all_students - isrdata %>%
  filter(!is.na(teacherid)) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many were excluded because the course wasn't actually eligible? 96
instructor_participated <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

instructor_participated - isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students") %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many were excluded because the instructor didn't adminster the correct survey? 31
eligible <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students") %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

eligible - isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester") %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many instructors did not complete survey 1? 0
eligible_correct <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester") %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

eligible_correct - isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many students did not consent? 116
teacher_complete <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()

teacher_complete - isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1
         & s1_consent == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# How many students did not complete survey 1? 304
student_consent <- isrdata %>%
  filter(!is.na(teacherid)) %>%
  filter(t1_consent == 1
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t1_consent == 1
         & t1_complete == 1
         & s1_consent == 1) %>%
  dplyr::select(id) %>%
  unique() %>%
  count() %>%
  pull()
